### 2018-07-05 Thesis/Parts/Background/Background.tex
updates
pc-re-read

\subheader{Neural Network}
A complete in-depth background into neural networks is beyond the scope of this thesis, therefore we will only describe the simplified working of a neural network and its basic application in IR.
Neural Networks are multi-layered computational software, based on matrix transformations, which tries to map input values to output values. This mapping uses a pre-defined amount of layers that modify the values through matrix transformations. Neural networks rely on training to create optimal values for these layers. Finding these optimal values is done iteratively, via a process known as "back-propogation". Back-propogation is done by starting at the output layer of the Neural Network and tracing error in the output back through the layers. This error information is used to improve the values of the layers. The values in the layers are only changed at training time; high-quality (training) data is therefore essential for neural networks. Neural networks can be applied to many different tasks, as long as there is sufficient data available to create a training and validation set. The training is the process of iteratively adjusting multiplication matrices or vectors to achieve the optimal result. Which is the result that is as close as possible to the given output for all input and output sets, without under performing on other sets\footnote{Having only good results for the training/validation set is know as overfitting. This means that the layer-values are fine-tuned to only perform (extremely) well on one set, while under performing on others}.\\Neural Networks are used, in the creation of word embeddings, to either predict words that may occur around a given word or predict a word given words that surround it. For example, given the sentence\\
\begin{center}
\textit{"The quick brown fox jumps over the lazy dog"}
\end{center}
the Neural Network can be trained to either predict the words around "jumps", which we refer to as context words (the, quick, brown, fox, over, the, lazy, dog). It can also predict the word "jumps", given the context words (the, quick, brown, fox, over, the, lazy, dog). This result in a matrix, a vector per word which is referred to as a word embedding. This embedding indicates "word relatedness" which, as mentioned earlier enables association (thus retrieval) with related(relevant) texts. It is to be noted that the word embeddings are only able to represent word relatedness to the other words trained in the same run. Due to the random initialization of the initial values of the layers in the neural networks, neural networks do not produce every run the same results. Therefore, word embeddings from different runs of the same neural network cannot be compared.

\subheader{Embedding}
An embedding is a distributed, numerical representation of text in a multi-dimensional vector space\footnote{The vector spaces of separately trained word embeddings differ, since each run the initial values of the neural network are randomly initialized. This means that the same word trained in two separate runs do not have to have the same embedding. However, they will have the same relationship to other words trained in their respective runs.} which can capture both the semantic and syntactic information\cite{mikolov2013distributed}. In the case of a word embeddings, the embeddings represent the words. These embeddings are created using machine learning models, which do not have the need for human interaction~\cite{lai2016generate}, they are so-called \textit{unsupervised learning algorithms}. Once trained, the embeddings can also be used to construct embeddings for collections of texts. For example, a word embedding can be used to create a sentence, paragraph, document or corpus embeddings. The usage of word embeddings has improved various Natural Language Processing areas such as named entity recognition, part-of-speech tagging, parsing, and semantic role labelling~\citet{luong2013better}.

@@

\subsubheader{TF-IDF}
Term Frequency * Inverted Document Frequency (TF-IDF) is a method that does not rely on a neural network and, therefore, does not require training. The TF-IDF score is the product of the term frequency in a text and the inverted document frequency of the same term in a corpus of texts. Both of which can be calculated in a variety of ways. Table~\ref{table:TFIDFScore} shows an example of the TF-IDF calculation. The multiplication of the TF and IDF is referred to as the TF-IDF score. The text, provided to the algorithm, is analyzed on word occurrences on corpus and document level, this results in a score per word. This score can be converted to a feature vector, in which the indexes of this vector represent the words and the value is the TF-IDF score. The size of this feature vector can optionally be controlled by hashing the words in the text, which can then be reduced to a given size. If this is not applied, the size of the feature vector is equal to the number of unique words in the corpus, also known as the (corpus) vocabulary.  The feature vectors produced by TF-IDF do not capture syntactic or semantic information about words, but capture information about word occurrences. Some applications of this technique limit the number of unique words in the text supplied to the TF-IDF algorithm, they do this by taking only a certain amount of top words, ordered on their occurrence. This reduces the amount of storage needed when hashing is not applied. It furthermore reduces the number of words which occur rarely. Due to the nature of TF-IDF, these rare words have a high score, cancelling out other more frequent words, while rarely occurring in the corpus.
\vspace{0.3in}\begin{table}[hbt]
\begin{center}
\begin{tabular}{|p{1in}|p{1in}|p{1in}|p{1.3in}|p{1in}|}
\hline
Term & Frequency (TF) & Document Frequency (DF) & Inverted Document Frequency (IDF) & TF-IDF score\\
\hline
Exponential & 4 & 15 & 0.824 & 3.296\\
\hline
Occurrence & 1 & 20 & 0.699 & 0.699\\
\hline
Multitude & 1 & 40 & 0.398 & 0.398\\
\hline
Abstract & 100 & 100 & 0 & 0\\
\hline
\end{tabular}
\caption{Example of TF-IDF score calculation.}\label{table:TFIDFScore}
\end{center}
\end{table}
\FloatBarrier
\end{jumpin}
As \citet{lai2016generate} state in their paper "How to generate a good word embedding?", that all embedding methods rely on the same hypothesis, \textit{words that occur in similar contexts have similar meanings}. They furthermore found that larger corpus' lead to better quality embeddings, but that the domain in which the embeddings are trained has more influence on this than the corpus size.

\subheader{Validation methods}
The results produced by the previously mentioned techniques have to be validated to determine their quality (in usage). The quality of the results can be validated through various validation metrics, one of these is the F1 score\footnote{The F1 score combines the precision score and the recall score in a single metric}. The embeddings can furthermore be validated through their performance on tasks such as word analogy, word similarity, categorization and embedding visualization. These tasks can be designed to produce a score that indicates the performance on a specific task. \citet{schnabel2015evaluation} found that a single validation metric cannot produce a representative result for other tasks. Embeddings that perform well on one task do not have to perform well on another task. As a result, findings about the performance of an embedding method are limited to the task on which they are tested. Their results cannot be generalized to state that the embeddings are overall "performing well". Validation tasks use either labeled or unlabeled data. Labeled data is data that is in some way marked so that the correct answer can be derived from it, while this is not possible with unlabelled data. The validation tasks can be divided into two groups, extrinsic evaluation, ones that use word embeddings as input for a downstream task and Intrinsic evaluation, which directly tests the relationships of the word embedding themselves. \citet{schnabel2015evaluation} note that the extrinsic evaluation may not be consistent with intrinsic evaluations since the performance on downstream tasks is not consistent across tasks.

@@

Since the word embeddings are created from a given text, these embeddings are bound to the text. All meaning embedded in the word embedding is derived from the original text. Because of this, embeddings can be "Domain-specific" meaning that it only knows words (or a specific word) in a certain context. This becomes most clear when faced with words that can have different meanings in different contexts. For this research, we categorize the embeddings into two categories, generic embeddings and domain-specific embeddings. The generic embeddings are trained on a collection of texts that use common English and contains a wide variety of topics. The domain-specific embeddings are trained on a collection of texts that uses jargon English (i.e., domain-specific terms) or is limited to a small number of topics. Given these terms, we regard the embeddings trained on the Wikipedia corpus\cite{lai2016generate, pennington2014glove, dai2015document, lau2016empirical, schnabel2015evaluation} as general, since Wikipedia uses common English and spans a wide range of topics. On the other hand, we regard the embeddings created by \citet{Truong2017Thesis} as Domain specific; these embeddings were created on academic articles, which use domain-specific terms and notations and only consists of academic texts, which contains less general/generic words compared to the Wikipedia corpus. Figure~\ref{figure:domainPlot} illustrates our corpus, marked in green, in comparison to the Wikipedia corpus, marked in blue. Our corpus is a collection of domain specific texts, written in academic language. A set that would be more specific would only focus on the individual domains or sub-domains.

@@

\caption{Illustration of generic texts and domain specific texts. Our research focusses on the green-marked area.}\label{figure:domainPlot}

@@

We defined domain specific earlier, a related term to this is "in-domain". The term domain-specific refers to a specific domain, I.E. biology or chemistry. In-domain refers to a relation between texts, this means that an article on biology is an in-domain text of another article on biology, but not on an article on chemistry. This means that two domain-specific texts do not necessarily share their domain, while an text A which is an in-domain text for text B means that text A and B share the same domain.

>>>>

\subheader{Neural Network}
A complete in-depth background into neural networks is beyond the scope of this thesis, therefore we will only describe the simplified working of a neural network and its basic application in IR.
Neural Networks are multi-layered computational software, based on matrix transformations, which tries to map input values to output values. This mapping uses a pre-defined amount of layers that modify the values through matrix transformations. Neural networks rely on training to create optimal values for these layers. Finding these optimal values is done iteratively, via a process known as "back-propagation". Back-propagation is done by starting at the output layer of the Neural Network and tracking error in the output back through the layers. This error information is used to improve the values of the layers. The values in the layers are only changed at training time; high-quality (training) data is, therefore, essential for neural networks. Neural networks can be applied to many different tasks, as long as there is sufficient data available to create a training and validation set. The training is the process of iteratively adjusting multiplication matrices or vectors values to achieve the optimal result. Which is the result that is as close as possible to the given output for all input and output sets, without underperforming on other sets\footnote{Having only good results for the training/validation set is known as overfitting. This means that the layer-values are fine-tuned to only perform (extremely) well on one set while underperforming on others}.\\Neural Networks are used, in the creation of word embeddings, to either predict words that may occur around a given word or predict a word given words that surround it. For example, given the sentence\\
\begin{center}
\textit{"The quick brown fox jumps over the lazy dog"}
\end{center}
the Neural Network can be trained to either predict the words around "jumps", which we refer to as context words (the, quick, brown, fox, over, the, lazy, dog). It can also predict the word "jumps", given the context words (the, quick, brown, fox, over, the, lazy, dog). This result in a matrix, a vector per word which is referred to as a word embedding. This embedding indicates "word relatedness" which, as mentioned earlier enables association (thus retrieval) with related(relevant) texts. It is to be noted that the word embeddings are only able to represent word relatedness to the other words trained in the same run. Due to the random initialization of the initial values of the layers in the neural networks, neural networks do not produce every run the same results. Therefore, word embeddings from different runs of the same neural network cannot be compared.

\subheader{Embedding}
An embedding is a distributed, numerical representation of text in a multi-dimensional vector space\footnote{The vector spaces of separately trained word embeddings differ, since each run the initial values of the neural network are randomly initialized. This means that the same word trained in two separate runs do not have to have the same embedding. However, they will have the same relationship to other words trained in their respective runs.} which can capture both the semantic and syntactic information\cite{mikolov2013distributed}. In the case of a word embeddings, the embeddings represent the words. These embeddings are created using machine learning models, which do not need human interaction~\cite{lai2016generate}, they are so-called \textit{unsupervised learning algorithms}. Once trained, the embeddings can also be used to construct embeddings for collections of texts. For example, a word embedding can be used to create a sentence, paragraph, document or corpus embeddings. The usage of word embeddings has improved various Natural Language Processing areas such as named entity recognition, part-of-speech tagging, parsing, and semantic role labelling~\citet{luong2013better}.

@@

\subsubheader{TF-IDF}
Term Frequency * Inverted Document Frequency (TF-IDF) is a method that does not rely on a neural network and, therefore, does not require training. The TF-IDF score is the product of the term frequency in a text and the inverted document frequency of the same term in a corpus of texts. Both of which can be calculated in a variety of ways. Table~\ref{table:TFIDFScore} shows an example of the TF-IDF calculation. The multiplication of the TF and IDF is referred to as the TF-IDF score. The text, provided to the algorithm, is analyzed on word occurrences on corpus and document level, this results in a score per word. This score can be converted to a feature vector, in which the indexes of this vector represent the words and the value is the TF-IDF score. The size of this feature vector can optionally be controlled by hashing the words in the text, which can then be reduced to a given size. If this is not applied, the size of the feature vector is equal to the number of unique words in the corpus, also known as the (corpus) vocabulary.  The feature vectors produced by TF-IDF do not capture syntactic or semantic information about words, but capture information about word occurrences. Some applications of this technique limit the number of unique words in the text supplied to the TF-IDF algorithm, they do this by taking only a certain amount of top words, ordered on their occurrence. This reduces the amount of storage needed when hashing is not applied. It furthermore reduces the number of words which occur rarely. Due to the nature of TF-IDF, these rare words have a high score, canceling out other more frequent words, while rarely occurring in the corpus.
\vspace{0.3in}\begin{table}[hbt]
\begin{center}
\begin{tabular}{|p{1in}|p{1in}|p{1in}|p{1.3in}|p{1in}|}
\hline
Term & Frequency (TF) & Document Frequency (DF) & Inverted Document Frequency (IDF) & TF-IDF score\\
\hline
Exponential & 4 & 15 & 0.824 & 3.296\\
\hline
Occurrence & 1 & 20 & 0.699 & 0.699\\
\hline
Multitude & 1 & 40 & 0.398 & 0.398\\
\hline
Abstract & 100 & 100 & 0 & 0\\
\hline
\end{tabular}
\caption{Example of TF-IDF score calculation.}\label{table:TFIDFScore}
\end{center}
\end{table}
\FloatBarrier
\end{jumpin}
As \citet{lai2016generate} state in their paper "How to generate a good word embedding?", that all embedding methods rely on the same hypothesis, \textit{words that occur in similar contexts have similar meanings}. They furthermore found that larger corpus' lead to better quality embeddings, but that the domain in which the embeddings are trained has more influence on this than the corpus size.

\subheader{Validation methods}
The results produced by the previously mentioned techniques have to be validated to determine their quality (in usage). The quality of the results can be validated through various validation metrics, one of these is the F1 score\footnote{The F1 score combines the precision score and the recall score in a single metric}. The embeddings can furthermore be validated through their performance on tasks such as word analogy, word similarity, categorization and embedding visualization. These tasks can be designed to produce a score that indicates the performance on a specific task. \citet{schnabel2015evaluation} found that a single validation metric cannot produce a representative result for other tasks. Embeddings that perform well on one task do not have to perform well on another task. As a result, the findings of the performance of an embedding method are limited to the task on which they are tested. Their results cannot be generalized to state that the embeddings are overall "performing well". Validation tasks use either labeled or unlabeled data. Labeled data is data that is in some way marked so that the correct answer can be derived from it, while this is not possible with unlabelled data. The validation tasks can be divided into two groups, extrinsic evaluation, ones that use word embeddings as input for a downstream task and Intrinsic evaluation, which directly tests the relationships of the word embedding themselves. \citet{schnabel2015evaluation} note that the extrinsic evaluation may not be consistent with intrinsic evaluations since the performance on downstream tasks is not consistent across tasks.

@@

Since the word embeddings are created from a given text, these embeddings are bound to the text. All meaning embedded in the word embedding is derived from the original text. Because of this, embeddings can be "Domain-specific" meaning that it only knows words (or a specific word) in a certain context. This becomes most clear when faced with words that can have different meanings in different contexts. For this research, we categorize the embeddings into two categories, generic embeddings and domain-specific embeddings. The generic embeddings are trained on a collection of texts that use common English and contains a wide variety of topics. The domain-specific embeddings are trained on a collection of texts that uses jargon English (i.e., domain-specific terms) or is limited to a small number of topics. Given these terms, we regard the embeddings trained on the Wikipedia corpus\cite{lai2016generate, pennington2014glove, dai2015document, lau2016empirical, schnabel2015evaluation} as general, since Wikipedia uses common English and spans a wide range of topics. On the other hand, we regard the embeddings created by \citet{Truong2017Thesis} as Domain specific; these embeddings were created on academic articles, which use domain-specific terms and notations and only consists of academic texts, which contains less general/generic words compared to the Wikipedia corpus. Figure~\ref{figure:domainPlot} illustrates our corpus, marked in green, in comparison to the Wikipedia corpus, marked in blue. Our corpus is a collection of domain-specific texts, written in academic language. A set that would be more specific would only focus on the individual domains or sub-domains.

@@

\caption{Illustration of generic texts and domain-specific texts. Our research focusses on the green-marked area.}\label{figure:domainPlot}

@@

We defined domain-specific earlier, a related term to this is "in-domain". The term domain-specific refers to a specific domain, I.E. biology or chemistry. In-domain refers to a relation between texts, this means that an article on biology is an in-domain text of another article on biology, but not on an article on chemistry. This means that two domain-specific texts do not necessarily share their domain, while a text A which is an in-domain text for text B means that text A and B share the same domain.

### 2018-07-05 Thesis/Parts/Data/Data.tex
pc-re-read

\caption{Word and token occurrences. The word counts pattern resembles a Pareto distribution, the tokenization reduced the amount of high frequency words.}\label{figure:wordTokenOccurrence}

@@

For this researched we used a (pre-made) tokenized dataset, created from the earlier described corpus, which reduces the total amount of words by 34\% (see table~\ref{table:corpusSize}). From this tokenized set, we created the embeddings and the TF-IDF feature vectors.

@@

For this research, we reused the word embeddings created by \citet{Truong2017Thesis}. These embeddings have a vector dimension of 300, which is an industry default. They have been trained on the entire Elsevier corpus, not limited to the subset we used for this research. To create higher-level embeddings, we average all component embeddings. Thus, to create article embeddings, we take the average of all normalized word embeddings for that article. Journal embeddings are created by taking the average of all normalized article embeddings. We use the average to combine multiple embeddings into one, because we determine the meaning of a larger text as the average meaning of all components. Since this meaning is represented as an embedding, we can average the embeddings to create the "average meaning". We have used multiple embedding optimizations for this research.

@@

Figure~\ref{figure:tfidfPerformance} shows the performance of title and abstract as median rank, and Figure~\ref{figure:tfidfMemory} shows the storage size in gigabyte, of the 1k/1K, 4K/4K, 7K/7K and 10K/10K TF-IDF configurations. This plot shows that, while the required storage size stagnate, the performance on title also quickly stagnates, together with the performance on the abstracts. The stagnation of the storage size is likely due to the fact that the terms which are added to the higher vocabulary sets occur less often than the other words (the cut-off is based on word occurrence, descending).Given this information, we have chosen to use the 10K/10K, 10K/5K and 5K/5K configurations to compare our embedding results to. Since these configurations can be kept in memory (storage size) while it is not likely that larger sets will give much improvement. The TF-IDF features are created on article level. We average the set of article feature vectors to create a journal feature vector, just as we did with the embeddings. 

@@

We procesessed the TF-IDF sets and the embedding sets via the same pipeline, using their common vector properties. This ensures comparable results, the pipeline is set-up as follows:

>>>>

\caption{Word and token occurrences. The word counts pattern resembles a Pareto distribution; the tokenization reduced the number of high-frequency words.}\label{figure:wordTokenOccurrence}

@@

For this research we used a (pre-made) tokenized dataset, created from the earlier described corpus, which reduces the total amount of words by 34\% (see table~\ref{table:corpusSize}). From this tokenized set, we created the embeddings and the TF-IDF feature vectors.

@@

For this research, we reused the word embeddings created by \citet{Truong2017Thesis}. These embeddings have a vector dimension of 300, which is an industry default. They have been trained on the entire Elsevier corpus, not limited to the subset we used for this research. To create higher-level embeddings, we average all component embeddings. Thus, to create article embeddings, we take the average of all normalized word embeddings for that article. Journal embeddings are created by taking the average of all normalized article embeddings. We use the average to combine multiple embeddings into one because we determine the meaning of a larger text as the average meaning of all components. Since this meaning is represented as an embedding, we can average the embeddings to create the "average meaning". We have used multiple embedding optimizations for this research.

@@

Figure~\ref{figure:tfidfPerformance} shows the performance of title and abstract as median rank, and Figure~\ref{figure:tfidfMemory} shows the storage size in gigabyte, of the 1k/1K, 4K/4K, 7K/7K and 10K/10K TF-IDF configurations. This plot shows that, while the required storage size stagnates, the performance on title also quickly stagnates, together with the performance on the abstracts. The stagnation of the storage size is likely due to the fact that the terms which are added to the higher vocabulary sets occur less often than the other words (the cut-off is based on word occurrence, descending). Given this information, we have chosen to use the 10K/10K, 10K/5K and 5K/5K configurations to compare our embedding results to. Since these configurations can be kept in memory (storage size) while it is not likely that larger sets will give much improvement. The TF-IDF features are created on article level. We average the set of article feature vectors to create a journal feature vector, just as we did with the embeddings. 

@@

We processed the TF-IDF sets and the embedding sets via the same pipeline, using their common vector properties. This ensures comparable results, the pipeline is set-up as follows:
